---
title: "p8105_hw3_qh2284"
author: "Qin Huang"
date: "2023-10-06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Problem 1

```{r}
# load the dataset
library(p8105.datasets)
data("instacart")
```

**Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations.**

```{r}
?instacart

user_num = instacart %>%
  distinct(user_id) %>%
  count()
```

The dataset contains `r nrow(instacart)` observations of `r user_num` unique users, where each row in the dataset is a product from an order.

There are `r ncol(instacart)` variables in this dataset:

-   `order_id`: order identifier, int type

-   `product_id`: product identifier, int type

-   `add_to_cart_order`: order where each product was added to cart, int type

-   `reordered`: 1 - reordered by this user (ordered in the past), 0 - not reorder, int type

-   `user_id`: customer identifier, int type

-   `eval_set`: which evaluation set this order belongs in (train), char type

-   `order_number`: the order sequence number for this user (1=first, n=nth), int type

-   `order_dow`: the day of the week on which the order was placed, int type

-   `order_hour_of_day`: the hour of the day on which the order was placed, int type

-   `days_since_prior_order`: days since the last order, capped at 30, NA if `order_number`=1, int type

-   `product_name`: name of the product, char type

-   `aisle_id`: aisle identifier, int type

-   `department_id`: department identifier, int type

-   `aisle`: the name of the aisle, char type

-   `department`: the name of the department, char type

**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
# count for the number of aisles
ai_num = instacart %>%
  distinct(aisle) %>%
  nrow()

# find the most items ordered
aisle_most = instacart %>%
  group_by(aisle) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(1) %>%
  pull(aisle)
```

There's total `r ai_num` aisles, among which `r aisle_most` are the most items ordered from.

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r}
# filter the aisles
more_than_1e4 = instacart %>%
  group_by(aisle) %>%
  summarise(n = n()) %>%
  filter(n > 10000) %>%
  arrange(desc(n))

# create the plot
plot_aisle = ggplot(more_than_1e4, aes(x = reorder(aisle, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +  
  labs(title = "Number of items ordered in each aisle",
       x = "Aisle",
       y = "Number of items ordered") +
  theme_minimal()

print(plot_aisle)
```

```{r}
# save the plot
ggsave("num of aisle.png", plot = plot_aisle)
```

We first group the aisles and count for each aisles. Then we filter the data. To make the plots more readable, we reorder the dataset to make it follow a descending order according to the number of aisles.

**Make a table showing the three most popular items in each of the aisles "baking ingredients", "dog food care", and "packaged vegetables fruits". Include the number of times each item is ordered in your table.**

```{r}
# create the table for top 3 popular items
popular_item = instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(ordered_num = n()) %>%
  arrange(aisle, desc(ordered_num)) %>%
  slice_head(n = 3)

print(popular_item)
```

According to the table, we could observe that the top 3 popular in "baking ingredients" are `Light Brown Sugar`, `Pure Baking Soda`, `Cane Sugar`. The top 3 popular in "dog food care" are `Snack Sticks Chicken & Rice Recipe Dog Treats`, `Organix Chicken & Brown Rice Recipe`, `Small Dog Biscuits`. The top 3 popular in "packaged vegetables fruits" are `Organic Baby Spinach`, `Organic Raspberries`, `Organic Blueberries`.

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
# first filter the table
table_mean_hour <- instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  # based on each day of the week
  group_by(product_name, order_dow) %>%
  # calculate the mean
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour)

print(table_mean_hour)
```

In this problem, we group the table based on the product name and the day of the week. Then we use summarise() to calculate the mean hour of orders for these two items.

------------------------------------------------------------------------

## Problem 2

**First, do some data cleaning:**

-   **format the data to use appropriate variable names;**

-   **focus on the "Overall Health" topic**

-   **include only responses from "Excellent" to "Poor"**

-   **organize responses as a factor taking levels ordered from "Poor" to "Excellent"**

```{r}
# load the dataset
data("brfss_smart2010")
```

```{r}
# clean the data
brfss_new = brfss_smart2010 %>%
  janitor::clean_names() %>%
  rename(state = locationabbr, loc_desc = locationdesc) %>%
  # focus on the “Overall Health” topic
  filter(topic == "Overall Health") %>%
  # include only responses from “Excellent” to “Poor”
  filter(!response %in% c("Yes","No","Yes, pregnancy-related")) %>%
  # organize responses as a factor taking levels ordered from “Poor” to “Excellent”
  mutate(response = factor(response, 
                           levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                           ordered = TRUE))
```

After cleaning, we would get a dataset with `r nrow(brfss_new)` observations and `r ncol(brfss_new)` variables.

**In 2002, which states were observed at 7 or more locations? What about in 2010?**

```{r}
# For 2002
states_2002 = brfss_new %>%
  filter(year == 2002) %>%
  group_by(state) %>%
  summarise(num_loc = n_distinct(loc_desc)) %>%
  filter(num_loc >= 7) %>%
  pull(state)

# For 2010
states_2010 = brfss_new %>%
  filter(year == 2010) %>%
  group_by(state) %>%
  summarise(num_loc = n_distinct(loc_desc)) %>%
  filter(num_loc >= 7) %>%
  pull(state)
```

In 2002, `r states_2002` could be observed at 7 or more locations, totally `r length(states_2002)` states.

In 2010, `r states_2010` could be observed at 7 or more locations, totally `r length(states_2010)` states.

**Construct a dataset that is limited to `Excellent` responses, and contains, year, state, and a variable that averages the `data_value` across locations within a state. Make a "spaghetti" plot of this average value over time within a state (that is, make a plot showing a line for each state across years -- the `geom_line` geometry and `group` aesthetic will help).**

```{r}
# construct only excellent dataset
brfss_excellent = brfss_new %>%
  filter(response == "Excellent") %>%
  group_by(year, state) %>%
  summarise(avg_data = mean(data_value, na.rm = T))

# create spaghetti plot
ggplot(brfss_excellent, aes(x = year, y = avg_data, group = state, color = state)) +
  geom_line() +
  labs(title = "Average Value Over Time by State",
       x = "Year",
       y = "Average Data Value") +
  theme_minimal()
```

Comment: We could observe that the average data value various among the year within different states. There aren't a tendency (increasing/decreasing) among the year, but fluctuate.

**Make a two-panel plot showing, for the years 2006, and 2010, distribution of `data_value` for responses ("Poor" to "Excellent") among locations in NY State.**

```{r}
# filter the year for only 2006 and 2010
data_ny = brfss_new %>%
  filter(year == 2006 | year == 2010) %>%
  filter(state == "NY") 

plot_ny = ggplot(data_ny, aes(x = response, y = data_value)) +
  geom_boxplot() +  
  facet_wrap(~ year, ncol = 2) +
  labs(title = "Distribution of Data Value by Response in NY State for 2006 and 2010",
       x = "Response",
       y = "Data Value") +
  theme_minimal()
  
plot_ny
```

Comment: For 2006 and 2010, we could observe similar distribution based on the response. They both have the largest average data value when the response is "very good", and the smallest average data value when the response is "poor". The rank of the mean is very good, good, excellent, fair, and poor. Also, we could observe that the "poor" group have the least variance in both year.

------------------------------------------------------------------------

## Problem 3

**Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).**

```{r}
# load dataset
accel = read_csv("./dataset/nhanes_accel.csv")
covar = read_csv("./dataset/nhanes_covar.csv", skip = 4)
```

```{r}
# merge the dataset by SEQN
combine_data = inner_join(covar, accel, by = "SEQN") %>%
  # filter year greater equal than 21
  filter(age >= 21) %>%
  # encode data with reasonable variable classes
  mutate(sex = if_else(sex == 1, "male", "female")) %>%
  mutate(education = case_when(
    education == 1 ~ "Less than high school",
    education == 2 ~ "High school equivalent",
    education == 3 ~ "More than high school"
  )) %>%
  mutate(sex = factor(sex, levels = c("male","female"), ordered = TRUE)) %>%
  mutate(education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school"), ordered = TRUE))  %>%
  # remove those with missing demographic data
  na.omit() %>%
  janitor::clean_names()
```

```{r}
# create a new table for tidy (the old table is for some plots)
combine_longer = combine_data %>%
  pivot_longer(min1:min1440,names_to = "time_mins", values_to = "mins_act") %>%
  na.omit()
```

After pivot_longer( ) , we could get a new dataset with `r nrow(combine_longer)` observations and `r ncol(combine_longer)` varaibles.

**Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.**

```{r}
# create the table
edu_sex_table <- combine_data %>%
  group_by(sex, education) %>%
  summarise(count = n()) %>%
  spread(key = sex, value = count)

edu_sex_table
```

We could observe that there are about half of the male and female have the education level of more than high school. We could simply observe the distribution of education within the male and female. There's not much difference between male and female at the education level of less than high school and more than high school. At the education level of high school equivalent, there are more males than females.

```{r}
# visualization
age_plot = ggplot(combine_data, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distributions by Sex and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal() +
  scale_fill_manual(values = c("male" = "lightblue", "female" = "lightpink"))

age_plot
```

Comment: According to the figure, we could observe that the average age of male and female who have education level of more than high school is smaller than other two groups. Within the same education level, female commonly have larger average age (except the education level of more than high school). And the male's age spread more than the female's (males' age has larger variance). The mean value of the age is similar between male and female at the education level of less than high school and more than high school. At the education level of high school equivalent, female have larger mean age than male.

**Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.**

```{r}
# calculate the total activity
total_ac = combine_longer %>%
  group_by(seqn,sex,age,bmi,education) %>%
  summarise(total = sum(mins_act)) 

act_plot = ggplot(total_ac, aes(x = age, y = total, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ education) +
  labs(title = "Total Activity by Age, Sex, and Education Level",
       x = "Age",
       y = "Total Activity") +
  theme_minimal() +
  scale_color_manual(values = c("male" = "blue", "female" = "brown"))

act_plot
```

Comment: The tendency of the total activity among age is decreasing in three education levels, which means, after about 40 years old, the more the age, the smaller the total activity value. At the education level of less than high school, male have more activity than female. At the education level of high school equivalent and more than high school, female have more activity than male. The range of the total activity for male and female at different education level are similar.

**Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.**

```{r}
# to get rid of the "min" in time for readability in x-axis
time_new = combine_longer %>%
  mutate(time_mins = as.numeric(gsub("min", "", time_mins))) %>%
  arrange(time_mins)

# create the plot
plot_24h = ggplot(time_new, aes(x = time_mins, y = mins_act, color = sex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ education, nrow = 2) +
  geom_smooth(se = FALSE) +
  labs(title = "24-hour Activity Time Courses for Each Education Level and Sex",
       x = "Time",
       y = "Activity")+
  theme_minimal()

plot_24h
```

Comment: We could observe that the smooth lines for male and female are almost the same, which indicates that there are not significant difference for their activity between male and female. And the trends in three education level are similar, indicating there are no significant difference for activity among education levels.
